Sparse Function-space Representation \\ of Neural Networks

### 2023-05-13 paper/main.tex
update methods

\subsection{Neural network to GP prior} 
\label{sec:nn2gp}
%
A Gaussian process characterizes a distribution over functions in terms of the first two moments. The key quantity of a Gaussian process is the prior covariance function $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior. If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx') = \delta^{-1} \phi(\vx) \phi(\vx')^\T$ see \citep{rasmussen2006gaussian} for details. We can approximate the feature map of the neural network  $f_\weights(\mathbf{x})$  by using the Jacobian at $\weights^*$ thus our GP prior becomes:
\begin{align} \label{eq-laplace-approx-function-space}
 g(\vx) \sim \Norm \left( \mu(\vx), k(\vx, \vx') \right) \quad
  \mu(\vx)
  = \vzeros \quad
  k(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\Jac{\weights^*}{\mathbf{x}'}^\top,
\end{align}
where the kernel is the Neural Tangent Kernel (NTK) with Jacobian given by $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$. Alternatively, we could have followed the derivation in \citet{immerImprovingPredictionsBayesian2021} where we perform a local linearization of the neural network prediction function $f^{\textrm{lin}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} (\vw - \vw^*)$. The difference between our GP prior mean function and \cref{eq-laplace-approx-function-space} is that we use a different mean prior function coming from $f^{\textrm{lin}}_{\vw^*}(\vx; \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^*$, essentially they use the difference from the linear model approximation and the neural network function, which does not seem useful as a prior mean function, so instead we use $ \mu(\vx_i) = \vzeros$ as is common in GP literature and derived from a feature map conversion.

As stated in \citet[Lemma~1]{csato2002sparse} the posterior process mean and covariance function of a GP prior combined with a likelihood function and $\dataset$ can be stated as:
\begin{equation}  \label{eq:gp_pred}
 \myexpect_{p(f\mid\vy)}[f_i] = \vk_{\vx i}^\top \valpha , \quad
\textrm{Var}_{p(f\mid\vy)}[f_i] = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
where the $ij$th entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$th element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(f_i \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(f_i \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.

>>>>

\subsection{Neural network to GP prior} 
\label{sec:nn2gp}
%
A Gaussian process characterizes a distribution over functions in terms of the first two moments. The key quantity of a Gaussian process is the prior covariance function (or kernel function) $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior. If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx') = \delta^{-1} \phi(\vx) \phi(\vx')^\T$ see \citep{rasmussen2006gaussian} for details.\todo{this part needs to be rewritten, starts talking about a general function parameterisation and then jumps directly to the neural network} 
We can approximate the feature map of the neural network  $f_\weights(\mathbf{x})$  by using the Jacobian at $\weights^*$ thus our GP prior becomes:
\begin{align} \label{eq-laplace-approx-function-space}
 g(\vx) \sim \Norm \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad
  \mu(\vx)
  = \vzeros \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\Jac{\weights^*}{\mathbf{x}'}^\top,
\end{align}
where the kernel is the Neural Tangent Kernel (NTK) \citep{jacot2018neural} with Jacobian $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$. Alternatively, we could have followed the derivation in \citet{immerImprovingPredictionsBayesian2021} where they perform a local linearization of the neural network  $f^{\textrm{lin}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} \, (\vw - \vw^*)$. Their mean prior function differ from ours, \cref{eq-laplace-approx-function-space}, as theirs comes from setting  $  \mu(\vx) = f^{\textrm{lin}}_{\vw_*}(\vx; \vw = \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^*$, \ie the difference between the linear model approximation and the neural network function at $\vx$. %which does not seem useful as a prior mean function, 
On the other hand, we set $\mu(\vx) = \vzeros$, as common in GP literature and derived from a feature map conversion. \todo{derived what?}

As stated in \citet[Lemma~1]{csato2002sparse}, the posterior process mean and covariance functions of a GP prior, combined with a likelihood function and $\dataset$ can be stated as:
\begin{equation}  \label{eq:gp_pred}
 \myexpect_{p(f_i \mid\vy)}[f_i] = \vk_{\vx i}^\top \valpha , \quad
\textrm{Var}_{p(f\mid\vy)}[f_i] = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
where $f_i = f_\weights(\vx_i)$, the $ij$th entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$th element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(f_i \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(f_i \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.
